\section{Introduction}

The discovery of meaningful patterns in temporal data streams has become increasingly important across numerous domains, from renewable energy systems \cite{qashou2022mining} to healthcare analytics \cite{yu2022data}. When such data is distributed across networked systems, the challenge of efficiently mining high-utility patterns while respecting communication constraints becomes particularly compelling. Recent work has explored various approaches to distributed pattern mining \cite{ding2022prefixpruningbased}, but fundamental questions about the theoretical limits and necessary conditions for efficient algorithms remain open.

In this paper, we study the theoretical foundations of distributed high-utility pattern mining over temporal data streams. Given a pattern space $P$ and utility function $U:P \to \mathbb{R}^+$, we seek to characterize precisely when efficient distributed algorithms exist for discovering patterns whose utility exceeds a given threshold. The key challenge lies in achieving near-optimal solutions while minimizing communication overhead across the network.

Traditional approaches to high-utility pattern mining have focused primarily on centralized settings \cite{wang2020improved}, where all data is available at a single node. However, modern applications increasingly involve distributed data sources that communicate over resource-constrained networks. This motivates our investigation of the fundamental properties that utility functions must satisfy to enable efficient distributed algorithms.

Our main contribution is a complete characterization of utility functions that admit efficient distributed mining algorithms. Specifically, we prove that convergence to within $\epsilon$ of the global optimum in $O(\frac{d_G}{\epsilon^2}\log|P|)$ rounds is possible if and only if the utility function satisfies an $\alpha$-smoothness property: $|U(p_1) - U(p_2)| \leq \alpha d_H(p_1,p_2)$ for all patterns $p_1,p_2 \in P$, where $d_H$ denotes Hamming distance. This characterization provides both necessary and sufficient conditions, settling a fundamental question in distributed pattern mining.

The rest of this paper is organized as follows. Section 2 formally defines our model and presents preliminary results. Section 3 introduces the $\alpha$-smoothness property and proves its necessity. Section 4 presents our distributed algorithm and establishes its convergence guarantees. Section 5 provides experimental validation, and Section 6 concludes with directions for future work.